---
title: "HW4"
author: "Patrick Gardocki"
date: "2023-11-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# 3.6: 28
$x^2y''-3xy'+4y=0; \ y(1)=5; \ y'1)=2$

$y'(x)=mx^{m-1}; \ y''(x)=m(m-1)x^{m-2}$

$(m(m-1)-3m+4)x^m=0 \\ (m^2-4m+4)x^m=0 \\ (m-2)^2=0 \\ \therefore m_1=m_2=2 \\ y_c(x)=c_1x^2+c_2x^2lnx$

$y'(x)=2c_1x+2c_2xlnx+c_2x \\ y(1)=5=c_1 \\ y'(1)=2=2c_1+c_2$

$c_1=5 \\ c_2=-8$

$y(x)=5x^2-8x^2lnx$


``` {r 1, echo=FALSE, out.width="50%"}
# Load the ggplot2 package
library(ggplot2)

# Create a sequence of x values
x <- seq(0.01, 5, by = 0.01)

# Calculate y values based on the function
y <- 5 * x^2 - 8 * x^2 * log(x)

# Create a data frame with x and y
data <- data.frame(x, y)

# Create the plot
plot <- ggplot(data, aes(x, y)) +
  geom_line() +
  labs(x = "x", y = "y") +
  ggtitle("Plot of y(x) = 5x^2 - 8x^2 * ln(x)")

# Display the plot
print(plot)

```

# 3.8: 41

$\frac{d^2x}{dt^2}+4x=-5sin2t+3cos2t; \ x(0)=-; \ x'(0)=1$

$m^2+4=0 \rightarrow m_1=-2i, \ m_2=2i$

$x_c(t)=c_1cos2t+c_2sin2t$

$x_p(t)=Atcos2t+Btsin2t \\ x'_p(t)=(A+2Bt)cos2t+(-2A+Bt)sin2t \\ x''_p(t)=(-4At+4B)cos2t + (-4A-4Bt)sin2t$

$4Bcos2t - 4Asin2t=3cos2t-5sin2t \\ 4B=3 \rightarrow B=\frac{3}{4}\\ -4A=-5 \rightarrow A= \frac{5}{4}$

$x_p(t)=\frac{5}{4}tcos2t + \frac{3}{4}tsin2t$

$x(t)=c_1cos2t+c_2sin2t+ \frac{5}{4}tcos2t + \frac{3}{4}tsin2t$

Given: $x(0)=-; \ x'(0)=1$

$x(t)=-cos2t-\frac{1}{8}sin2t+ \frac{5}{4}tcos2t + \frac{3}{4}tsin2t$

# 3.9: 12

$y''+\lambda y=0; \ y'(0)=0; \ y(\pi/4)=0$

For $\lambda =0$, $y(x)=c_1x+c_2, \ y'(0)=y(\pi/4)=0 \rightarrow c_1=c_2=0$

$y(x)=0$ For $\lambda=0$, no nontrivial solutions, and is not eigenvalue.

For $\lambda <0$, $\lambda = -\alpha^2; \ y''-\alpha^2y=0; \ m^2-\alpha^2=0 \rightarrow m_1=-\alpha, \ m_2=\alpha$

$y(x) = c_1cosh\alpha x+c_2sinh \alpha \rightarrow c_1=c_2=0 \rightarrow y(x)=0$

For $\lambda<0$, no nontrivial solutions, and is not eigenvalue.


For $\lambda >0$, $\lambda = \alpha^2; \ y''+\alpha^2y=0; \ m^2+\alpha^2=0 \rightarrow m_1=-i\alpha, \ m_2=i\alpha$

$c_2=0; \ c_1cos\frac{\alpha \pi}{4}=0$

For $c_1 \ne0$, nontrivial solutions. $cos\frac{\alpha \pi}{4}=0 \rightarrow \alpha=2+4n$

$\lambda_n=(2+4n)^2$

$y_n(x)=sin(2+4n)x$


# 5.1: 8
$\frac{1-x}{2+x}=\frac{3-2-x}{2-x}=-1+\frac{3}{2+x}=-1+(\frac{3}{2})(\frac{1}{1+\frac{x}{2}})=\frac{1}{2}-\frac{3x}{4}+\frac{3x^2}{8}-\frac{3x^3}{16}+...$


$|\frac{x}{2}|<1 \rightarrow$ Interval of Convergence $(-2,2)$

# 5.1: 19

$y=\sum c_nx^n$

$y''-2xy'+y=0=\sum n(n-1)(c_nx^{n-2})-2\sum nc_nx^n+\sum c_nx^n \\= \sum (k+2)(k+1)c_{k+2}x^k-s\sum kc_kx^k + \sum c_kx^k \\ = 2c_2+c_0+\sum ((k+2)(k+1)(k+2)-(2k-1)c_k)x^k$

$2c_2=-\frac{1}{2}c_0$

$c_{k+2}=\frac{2k-1}{(k+2)(k+1)}c_k$

For $c_0=1; \ c_1=0$

$c_2=-\frac{1}{2}; \ c_3=c_5=c_7=0; \ c_4=-\frac{1}{8}; \ c_6=-\frac{7}{240}$

For $c_0=0; \ c_1=1$

$c_2=c_4=c_6=0; \ c_3=\frac{1}{6}; \ c_5=\frac{1}{24}; \ c_7=\frac{1}{112}$

$\therefore y_1= 1-\frac{x^2}{2}-\frac{x^4}{8}-\frac{7x^6}{240}; \ y_2= x+ \frac{x^3}{6} + \frac{x^5}{24} + \frac{x^7}{112}$

# 5.2: 6

$x^2(x-5)^2y''+4xy'+(x^2-25)y=0 \rightarrow$

$P(x)=\frac{4x}{x^2(x-5^2)}; \ Q(x)=\frac{(x^2-25)}{x^2(x-5)^2}$

For $x_0=0$, $p(x)=(x-0)\frac{4x}{x^2(x-5)^2}; \ q(x)=(x-0)^2\frac{(x^2-25)}{x^2(x-5)^2}. \ \therefore x=0$ is regular singular point.

For $x_0=5$, $p(x)=(x-5)\frac{4x}{x^2(x-5)^2}; \ q(x)=(x-5)^2\frac{(x^2-25)}{x^2(x-5)^2}. \ \therefore x=5$ is irregular singular point.

# 5.2: 19

$3xy''+ (2-x)y'-y=0$

$3\sum (n+r-1)(n+r)c_nx^{n+r-1} + 2\sum (n+r)c_nx^{n+r-1}-\sum (n+r)c_nx^{n+r}-\sum c_nx^{n+r}=0$

$(3r^2-r)c_0x^{r-1}+\sum_{n=1}(3(n+r)^2-(n+r))c_nx^{n+r-1}-\sum (n+r+1)c_nx^{n+r}=0$

Sub m for n-1: $(3r^2-r)c_0x^{r-1}+\sum_{m=0}(((3(m+1+r)^2-(m+1+r))c_{m+1}- (m+r+1)c_m)x^{m+r}=0$

$c_{m+1}=\frac{c_m}{3(m+r+1)-1}$

Given: $3r^2-r=r(3r-1)=0;$ roots are: $r_1=0, \ r_2 = \frac{1}{3}$

For $r_1=0$

$c_1=\frac{1}{2}c_0; \ c_2=\frac{1}{10}c_0; \ c_3=\frac{1}{80}c_0 \ \therefore y_1=1+\frac{1}{2}x+\frac{1}{10}x^2 + \frac{1}{80}x^3+...$

For $r_2=\frac{1}{3}$

$c_1=\frac{1}{3}c_0; \ c_2=\frac{1}{18}c_0; \ c_3=\frac{1}{162}c_0 \ \therefore y_1=x^{1/3}(1+\frac{1}{3}x+\frac{1}{18}x^2 + \frac{1}{162}x^3+...)$

$y=c_1(1+\frac{1}{2}x+\frac{1}{10}x^2 + \frac{1}{80}x^3+...)+ c_2x^{1/3}(1+\frac{1}{3}x+\frac{1}{18}x^2 + \frac{1}{162}x^3+...)$


# 10.2: 22

$det(A-\lambda I)=0$

$\begin{vmatrix} -6-\lambda & 5 \\ -5 & 4-\lambda \end{vmatrix}=0$

$-24-4\lambda + 6\lambda + \lambda^2+25=0=\lambda^2+2\lambda +1 \rightarrow \lambda=-1$

$\begin{vmatrix} -6+1 & 5 \\ -5 & 4+1 \end{vmatrix}=0 \rightarrow \begin{vmatrix} 1 & -1 \\ 0 & 0 \end{vmatrix}$

$x_1-x_2=0 \rightarrow x_1=x_2$

$K=\begin{bmatrix} 1 \\ 1\end{bmatrix}$

$\begin{vmatrix} -6+1 & 5 \\ -5 & 4+1 \end{vmatrix}=\begin{bmatrix} 1 \\ 1\end{bmatrix} \rightarrow \begin{vmatrix} 1 & -1 \\ 0 & 0 \end{vmatrix}=\begin{bmatrix} \frac{-1}{5} \\ 0\end{bmatrix}$

$x_1-x_2=\frac{-1}{5}; \ x_1=\frac{-1}{5}+x_2$

$K_2=\begin{bmatrix} 0 \\ \frac{1}{5}\end{bmatrix}$

$X=c_1(\begin{bmatrix} 1 \\ 1\end{bmatrix}e^{-t} + c_2(\begin{bmatrix} 1 \\ 1\end{bmatrix}t+\begin{bmatrix} 0 \\ \frac{1}{5}\end{bmatrix} )e^{-t}= \begin{bmatrix} c_1e^{-t}+ c_2te^{-t} \\ c_1e^{-t}+c_2(t+\frac{1}{5}e^{-t}) \end{bmatrix}$

# 10.2: 38

$det(A-\lambda I)=0$

$\begin{vmatrix} 4-\lambda & 5 \\ -2 & 6-\lambda \end{vmatrix}=0$

$24-4\lambda - 6\lambda + \lambda^2+10=0=\lambda^2-10\lambda +34 \rightarrow \lambda=5 \pm3i$

$\begin{vmatrix} 4-(5+3i_) & 5 \\ -2 & 6-(5+3i) \end{vmatrix}=0 \rightarrow \begin{vmatrix} 1 & \frac{-1}{2}+\frac{3}{2}i \\ 0 & 0 \end{vmatrix}$

$x_1=-(\frac{-1}{2}+\frac{3}{2}i)x_2$

$K=\begin{bmatrix} 1-3i \\ 2\end{bmatrix}; \ Re(K)=K=\begin{bmatrix} 1 \\ 2\end{bmatrix}; \ Im(K)=K=\begin{bmatrix} -3 \\ 0\end{bmatrix}$

$X=c_1(\begin{bmatrix} 1 \\ 2\end{bmatrix}cos3t-\begin{bmatrix} -3 \\ 0\end{bmatrix}sin3t) + c_2(\begin{bmatrix} -3 \\ 0\end{bmatrix}cos3t+\begin{bmatrix} 1 \\ 2\end{bmatrix}sin3t )= c_1\begin{bmatrix} cos3t+3sin3t \\ 2cos3t\end{bmatrix}e^{5t}+c_2\begin{bmatrix} sin3t-3cos3t \\ 2sin3t\end{bmatrix}e^{5t}$




# 10.4: 14 by using variation of parameters using the matrix exponential formula

$\frac{dx}{dt}=2x-y; \\ \frac{dy}{dt}=3x-2y+4t$

$\begin{vmatrix} 2-\lambda & -1 \\ 3 & -2-\lambda \end{vmatrix} \rightarrow \lambda^2-1=0; \ \lambda=-1,1$

For,$\lambda=-1$, eigenvector $=\begin{bmatrix} 1 \\ 1\end{bmatrix}$

For,$\lambda=1$, eigenvector $=\begin{bmatrix} 1 \\ 3\end{bmatrix}$


$x_c(t)=c_1\begin{bmatrix} 1 \\ 1\end{bmatrix}e^t+c_2\begin{bmatrix} 1 \\ 3\end{bmatrix}e^{-t}$

$\phi =\begin{bmatrix} e^t & e^{-t} \\ e^t& 3e^{-t}\end{bmatrix}; \ \phi^{-1}=\begin{bmatrix} \frac{3}{2}e^{-t} & \frac{-1}{2}e^{-t} \\ \frac{1}{2}e^t& \frac{1}{2}e^{t}\end{bmatrix}$

$X_p=\phi \int\phi^{-1}Fdt=\begin{bmatrix} e^t & e^{-t} \\ e^t& 3e^{-t}\end{bmatrix} \int \begin{bmatrix} \frac{3}{2}e^{-t} & \frac{-1}{2}e^{-t} \\ \frac{1}{2}e^t& \frac{1}{2}e^{t}\end{bmatrix} \begin{bmatrix} 0 \\ 4 \end{bmatrix}= \begin{bmatrix} e^t & e^{-t} \\ e^t& 3e^{-t}\end{bmatrix} \begin{bmatrix} 2te^{-t}+2e^{-t} \\ 2te^t-2e^{t}\end{bmatrix}=\begin{bmatrix} 4 \\ 8\end{bmatrix}t+ \begin{bmatrix} 0 \\ -4\end{bmatrix}$





